acquisition of TEs

### 2021-03-22 Analyze_TEs.Rmd
updated observed data analysis

  #Defining vocabulary-based balance score
  mutate(balance_vocab = TotalNW_NonDom/TotalNW_Dom) %>%
  #Defining proportion of TE (with reference to child's own total vocab size)
  mutate(propTE = number_of_te*2/word_vocab) %>% 
  #Defining proportion of TE (with reference to CDI)
  mutate(propTE_CDI = number_of_te/672)

@@

@@

## Merge the two percentile dataframe
ws_percentile <- merge(english_ws_percentile, french_ws_percentile, 
                       by = "age") %>%
  # compute estimate total vocab by adding the two languages
  mutate(Total_90percentile = EngWS_90percentile + FrWS_90percentile) %>%
  # keep only necessary columns
  select(age, EngWS_90percentile, FrWS_90percentile, Total_90percentile) %>%
  # rename age to prepare for the next step (i.e., merging with keepers_ws_TE
  rename(age_months_percentile = age)
```

# Adding percentile info to keepers_ws_TE dataframe & 
## Creating a new propTE based on the percentile
```{r}
# merge the dataframes
keepers_ws_TE <- keepers_ws_TE %>%
  # as the upper age limit of the CDI-WS is 30m, create a new age variable which replaces all >30m as 30m for merging with the percentile dataframe
  # this means the percentile should be the same for all >30m
  mutate(age_months_percentile = ifelse(age_months_binned > 30, 30, age_months_binned)) %>%
  # merge with the percentile data frame
  left_join(ws_percentile, by = "age_months_percentile") %>%
  # calculate TE in relation to CDI percentile data
  mutate(propTE_percentile = number_of_te*2/Total_90percentile)
  # %>% select(baby_id, age_months_binned, age_months_percentile, word_vocab, Total_90percentile,
  #       propTE, propTE_CDI, propTE_percentile)

@@

## Create predicted TEs for simulation
```{r}
keepers_ws_TE_simulation <- keepers_ws_TE %>%
  # Predicted propTE (Simulation)
  mutate(propDomVocab = TotalNW_Dom / 672, # Proportion of Dom/NonDom vocab
         propNonDomVocab = TotalNW_NonDom / 672, 
         Predicted_propTE = propDomVocab * propNonDomVocab, # Predicted proportion of TE
         Diff = propTE - Predicted_propTE) %>% # Difference between Predicted and Observed propTE
  # Predicted raw TE (with reference to CDI)
  mutate(Predicted_TotalnumberTEs_CDI = (TotalNW_Dom * TotalNW_NonDom)/672) %>%
  # Predicted raw TE (with reference to CDI percentile)
  mutate(Predicted_TotalnumberTEs_percentile = TotalNW_Dom * TotalNW_NonDom/(Total_90percentile/2)) %>%
  # Create two groups based on observed TE
  mutate(TEgroup = ifelse(number_of_te < 100, "Observed TE less than 100", "Observed TE more than 100"))

  #select(baby_id, age_months_binned, age_months_percentile, word_vocab, Total_90percentile,
  #       propTE, propTE_CDI, propTE_percentile, Predicted_propTE, Diff, 
  #       Predicted_TotalnumberTEs_CDI, Predicted_TotalnumberTEs_percentile)

@@

###### Linear regressions of predicted TE models ###### 
```{r}

### Model 1: Predicted_TotalnumberTEs_percentile
Percentile_model <- keepers_ws_TE_simulation %>%
  lm(number_of_te*(Total_90percentile/2) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
     data = .)

### Model 2: Predicted_TotalnumberTEs_CDI
CDI_model <- keepers_ws_TE_simulation %>%
  lm(number_of_te * 672 ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
     data = .)

anova(Percentile_model, CDI_model)
```

###### Theoretical distribution of TEs (Simulation) ###### 
```{r}
###Prediction 1: Predicted raw TE = (TotalNW_Dom*TotalNW_NonDom)/Total CDI
# Ploting Y = observed rawTE  & X = Predicted raw TE

keepers_ws_TE_simulation %>%
  ggplot(aes(x = Predicted_TotalnumberTEs_CDI, y = number_of_te)) +
  facet_grid(. ~TEgroup) +
  labs(x = "Predicted total number of TEs (based on CDI)", y = "Observed total number of TEs",
       title = "Observed vs. Predicted total number of TEs") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_smooth(method = "lm", se = F) +
  geom_point(shape = 23) +
  geom_abline(linetype = 2, alpha = 0.5) +
  xlim(-10,430) +
  ylim(-10,430)

# Kolmogorov-Smirnov test to compare the observed and predicted TE
stats::ks.test(keepers_ws_TE_simulation$number_of_te,
               keepers_ws_TE_simulation$Predicted_TotalnumberTEs_CDI)

###Prediction 2: Predicted raw TE = TotalNW_Dom * TotalNW_NonDom/(Total_90percentile/2)
# Ploting Y = observed rawTE  & X = Predicted raw TE

TEprediction_percentile <- keepers_ws_TE_simulation %>%
  ggplot(aes(x = Predicted_TotalnumberTEs_percentile, y = number_of_te)) +
  facet_grid(. ~TEgroup) +
  labs(x = "Predicted total number of TEs (based on percentile)", y = "Observed total number of TEs",
       title = "Observed vs. Predicted total number of TEs") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_smooth(method = "lm", se = F) +
  geom_point(shape = 23) +
  geom_abline(linetype = 2, alpha = 0.5) +
  xlim(-10,430) +
  ylim(-10,430)

ggsave(here("Output/TEprediction_percentile.png"), TEprediction_percentile)

# Kolmogorov-Smirnov test to compare the observed and predicted TE
stats::ks.test(keepers_ws_TE_simulation$number_of_te,
               keepers_ws_TE_simulation$Predicted_TotalnumberTEs_percentile)

keepers_ws_TE_simulation_LessThan10 <- keepers_ws_TE_simulation %>%
  filter(TEgroup == "Observed TE less than 100")

stats::ks.test(keepers_ws_TE_simulation_LessThan10$number_of_te,
               keepers_ws_TE_simulation_LessThan10$Predicted_TotalnumberTEs_percentile)

keepers_ws_TE_simulation_MoreThan10 <- keepers_ws_TE_simulation %>%
  filter(TEgroup == "Observed TE more than 100")

stats::ks.test(keepers_ws_TE_simulation_MoreThan10$number_of_te,
               keepers_ws_TE_simulation_MoreThan10$Predicted_TotalnumberTEs_percentile)

### Plotting the 2 predictions together

TEprediction_percentile_CDI <- keepers_ws_TE_simulation %>%
  select(baby_id, number_of_te, Predicted_TotalnumberTEs_CDI, Predicted_TotalnumberTEs_percentile, TEgroup) %>%
  dplyr::rename("=Dom*NonDom/TotalCDI" = Predicted_TotalnumberTEs_CDI,
                "=Dom*NonDom/Percentile" = Predicted_TotalnumberTEs_percentile) %>%
  pivot_longer(-c(baby_id, number_of_te, TEgroup), 
               names_to = "Predicted_TotalnumberTEs", values_to = "PredictedNumber") %>%
  ggplot(aes(x = PredictedNumber, y = number_of_te, color = Predicted_TotalnumberTEs)) +
  facet_grid(. ~TEgroup) +
  labs(x = "Predicted total number of TEs", 
       y = "Observed total number of TEs",
       title = 
       "Observed vs. Predicted total number of TEs") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_smooth(method = "lm", se = F) +
  geom_point(shape = 23) +
  geom_abline(linetype = 2, alpha = 0.5) +
  xlim(-10,500) +
  ylim(-10,500) +
  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") +
  guides(colour = guide_legend(nrow = 1))

ggsave(here("Output/TEprediction_percentile&CDI.png"), TEprediction_percentile_CDI)

#Less Than 100 observed TEs
TEprediction_percentile_CDI_LessThan100 <- keepers_ws_TE_simulation %>%
  filter(TEgroup == "Observed TE less than 100") %>%
  select(baby_id, number_of_te, Predicted_TotalnumberTEs_CDI, Predicted_TotalnumberTEs_percentile) %>%
  dplyr::rename("=Dom*NonDom/TotalCDI" = Predicted_TotalnumberTEs_CDI,
                "=Dom*NonDom/Percentile" = Predicted_TotalnumberTEs_percentile) %>%
  pivot_longer(-c(baby_id, number_of_te), 
               names_to = "Predicted_TotalnumberTEs", values_to = "PredictedNumber") %>%
  ggplot(aes(x = PredictedNumber, y = number_of_te, color = Predicted_TotalnumberTEs)) +
  labs(x = "Predicted total number of TEs", 
       y = "Observed total number of TEs",
       title = 
       "Observed vs. Predicted total number of TEs \n (Less than 100 observed TEs)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_smooth(method = "lm", se = F) +
  geom_point(shape = 23) +
  geom_abline(linetype = 2, alpha = 0.5) +
  xlim(-10,100) +
  ylim(-10,100) +
  theme(plot.title = element_text(size = 12, face = "bold"), 
        axis.text.x = element_text(angle = 90), 
        legend.position = "bottom", legend.text=element_text(size = 10)) +
  guides(colour = guide_legend(nrow = 2))

ggsave(here("Output/TEprediction_percentile&CDI_LessThan100.png"), TEprediction_percentile_CDI_LessThan100)

#Less Than 50 observed TEs
TEprediction_percentile_CDI_LessThan50 <- keepers_ws_TE_simulation %>%
  filter(number_of_te < 50) %>%
  select(baby_id, number_of_te, Predicted_TotalnumberTEs_CDI, Predicted_TotalnumberTEs_percentile) %>%
  dplyr::rename("=Dom*NonDom/TotalCDI" = Predicted_TotalnumberTEs_CDI,
                "=Dom*NonDom/Percentile" = Predicted_TotalnumberTEs_percentile) %>%
  pivot_longer(-c(baby_id, number_of_te), 
               names_to = "Predicted_TotalnumberTEs", values_to = "PredictedNumber") %>%
  ggplot(aes(x = PredictedNumber, y = number_of_te, color = Predicted_TotalnumberTEs)) +
  labs(x = "Predicted total number of TEs", 
       y = "Observed total number of TEs",
       title = 
       "Observed vs. Predicted total number of TEs \n (Less than 50 observed TEs)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_smooth(method = "lm", se = F) +
  geom_point(shape = 23) +
  geom_abline(linetype = 2, alpha = 0.5) +
  xlim(-10,50) +
  ylim(-10,50) +
  theme(plot.title = element_text(size = 12, face = "bold"),
        axis.text.x = element_text(angle = 90), 
        legend.position = "bottom", legend.box="vertical", legend.text=element_text(size = 10)) +
  guides(colour = guide_legend(nrow = 2))

ggsave(here("Output/TEprediction_percentile&CDI_LessThan50.png"), TEprediction_percentile_CDI_LessThan50)

# More than 100 Observed TEs
TEprediction_percentile_CDI_MoreThan100 <- keepers_ws_TE_simulation %>%
  filter(TEgroup == "Observed TE more than 100") %>%
  select(baby_id, number_of_te, Predicted_TotalnumberTEs_CDI, Predicted_TotalnumberTEs_percentile) %>%
  dplyr::rename("=Dom*NonDom/TotalCDI" = Predicted_TotalnumberTEs_CDI,
                "=Dom*NonDom/Percentile" = Predicted_TotalnumberTEs_percentile) %>%
  pivot_longer(-c(baby_id, number_of_te), 
               names_to = "Predicted_TotalnumberTEs", values_to = "PredictedNumber") %>%
  ggplot(aes(x = PredictedNumber, y = number_of_te, color = Predicted_TotalnumberTEs)) +
  labs(x = "Predicted total number of TEs", 
       y = "Observed total number of TEs",
       title = "Observed vs. Predicted total number of TEs \n (More than 100 observed TEs)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_smooth(method = "lm", se = F) +
  geom_point(shape = 23) +
  geom_abline(linetype = 2, alpha = 0.5) +
  xlim(-10,500) +
  ylim(-10,500) +
  theme(plot.title = element_text(size = 12, face = "bold"),
        axis.text.x = element_text(angle = 90), 
        legend.position = "bottom", legend.text=element_text(size=10)) +
  guides(colour = guide_legend(nrow = 2))

ggsave(here("Output/TEprediction_percentile&CDI_MoreThan100.png"), TEprediction_percentile_CDI_MoreThan100)

###Prediction 2: Predicted raw TE = TotalNW_NonDom
# Ploting Y = observed rawTE  & X = Predicted raw TE

>>>>

  #Defining vocabulary-based balance score
  #mutate(balance_vocab = TotalNW_NonDom/TotalNW_Dom) %>%
  mutate(balance_vocab = TotalNW_Dom/(TotalNW_Dom+TotalNW_NonDom)) %>%
  #Defining proportion of TE (with reference to child's own total vocab size)
  mutate(propTE = number_of_te*2/word_vocab) %>% 
  #Defining proportion of TE (with reference to CDI)
  mutate(propTE_CDI = number_of_te/672)

write.csv(keepers_ws_TE, "Data/keepers_ws_TE.csv", row.names = F)

@@

# Figure 6
```{r}
Observed_plot_inputs_TE <- keepers_ws_TE %>% 
  select(baby_id, ID_testdate, age_days, number_of_te, word_vocab, TotalNW_Dom, TotalNW_NonDom, balance_vocab) %>%
  pivot_longer(c(word_vocab, TotalNW_Dom, TotalNW_NonDom), names_to = "vocab_type", values_to = "number") %>%
  mutate(vocab_type = factor(vocab_type, levels=c("word_vocab", "TotalNW_Dom", "TotalNW_NonDom"))) %>%
  mutate(vocab_type = recode(vocab_type, 
                             word_vocab = "Panel A: Total vocabulary", 
                             TotalNW_Dom = "Panel B: Dominant vocabulary", 
                             TotalNW_NonDom = "Panel C: Non-dominant vocabulary")) %>%
  mutate(bin_B = case_when(balance_vocab < 0.6 ~ "0.5",
                           balance_vocab >= 0.6 & balance_vocab < 0.7 ~ "0.6",
                           balance_vocab >= 0.7 & balance_vocab < 0.8 ~ "0.7",
                           balance_vocab >= 0.8 & balance_vocab < 0.9 ~ "0.8",
                           TRUE ~ "0.9")) %>%
  ggplot(aes(x = number, y = number_of_te, color = bin_B)) +
  #geom_point() +
  stat_smooth(method = lm, se = F) +
  theme_minimal() + 
  facet_grid(. ~ vocab_type) +
  ylim(0, 400) +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Number of words", 
       y = "Total observed number of translation equivalents", 
       color = "Balance"
       #, title = "Number of translation equivalent by balance in relation to different input measures "
       ) +
  guides(colour = guide_legend(nrow = 1))

ggsave("Output/Observed_plot_inputs_TE.png", Observed_plot_inputs_TE,
       width = 10, height = 5)

# Figure 7
```{r}
plot_stacked_byV <- keepers_ws_TE %>%
  select(baby_id, age_months_percentile, balance_vocab, word_vocab, number_of_te, Dom_unique_words, NonDom_unique_words) %>%
  # Subset developmental level (i.e., Average_90percentile) by 3 levels 
  mutate(V_subset = case_when(age_months_percentile <= 22 ~ "18-22 months",
                              age_months_percentile > 22 & age_months_percentile <= 27 ~ "23-27 months",
                              TRUE ~ "28-33 months")) %>%
  #mutate(V_subset = case_when(age_months_percentile <= 21 ~ "18-21",
  #                            age_months_percentile > 21 & age_months_percentile <= 25 ~ "22-25",
  #                            age_months_percentile > 25 & age_months_percentile <= 29 ~ "26-29",
  #                            TRUE ~ "30-33")) %>%
  #mutate(bin_B = case_when(balance_vocab < 0.6 ~ "0.5",
  #                         balance_vocab >= 0.6 & balance_vocab < 0.7 ~ "0.6",
  #                         balance_vocab >= 0.7 & balance_vocab < 0.8 ~ "0.7",
  #                         balance_vocab >= 0.8 & balance_vocab < 0.9 ~ "0.8",
  #                         TRUE ~ "0.9")) %>%
  mutate(BalSubset = case_when(balance_vocab <= 0.65 ~ "Most Balanced",
                               balance_vocab >= 0.80 ~ "Least Balanced",
                               TRUE ~ "Medium Balanced")) %>%
  mutate(BalSubset = factor(BalSubset, levels=c("Most Balanced", "Medium Balanced", "Least Balanced"))) %>%
  arrange(desc(number_of_te)) %>%
  mutate(TE = number_of_te) %>%
  pivot_longer(-c(baby_id, age_months_percentile, balance_vocab, V_subset, BalSubset, word_vocab, TE), 
               names_to = "Type", values_to = "NumberOfWords") %>%
  mutate(Type = factor(Type, levels=c("NonDom_unique_words", "Dom_unique_words", "number_of_te"))) %>%
  mutate(Type = recode(Type, 
                       NonDom_unique_words = "Unique non-dominant word", 
                       Dom_unique_words = "Unique dominant word", 
                       number_of_te = "Translation equivalent")) %>%
  group_by(V_subset, BalSubset, Type) %>%
  summarise(n_words = mean(NumberOfWords)) %>%
  ggplot(aes(x = BalSubset, y = n_words, fill = Type)) +
  geom_bar(position="stack", stat="identity") +
  facet_grid(. ~ V_subset) +
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
  #scale_y_continuous(limits = c(0, 1200)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0) , 
        legend.position = "bottom",
        plot.title = element_text(hjust = 0.5)) + 
  labs(x = "Balance", 
       y = "Observed number of concepts", 
       title = "Vocabulary composition across different subset of developmental timepoints")

ggsave("Output/Observed_plot_stacked_byV.png", plot_stacked_byV,
       width = 10, height = 5)

## Mean proportion of TE
stacked_byV_wide <- keepers_ws_TE %>%
  select(baby_id, age_months_percentile, balance_vocab, word_vocab, concept_vocab, 
         propTE, number_of_te, Dom_unique_words, NonDom_unique_words) %>%
  # Subset developmental level (i.e., Average_90percentile) by 3 levels 
  mutate(V_subset = case_when(age_months_percentile <= 22 ~ "18-22 months",
                              age_months_percentile > 22 & age_months_percentile <= 27 ~ "23-27 months",
                              TRUE ~ "28-33 months")) %>%
  mutate(BalSubset = case_when(balance_vocab <= 0.65 ~ "Most Balanced",
                               balance_vocab >= 0.80 ~ "Least Balanced",
                               TRUE ~ "Medium Balanced"))

stacked_byV_wide %>%
  group_by(V_subset) %>%
  summarise(mean_propTE = mean(propTE, na.rm=T), 
            sd_propTE = sd(propTE, na.rm=T),
            min_propTE = min(propTE, na.rm=T),
            max_propTE = max(propTE, na.rm=T))

# Anova to compare the number of TEs among the 3 developmental groups in the stacked barchart
stacked_byV_wide %>%  
  group_by(V_subset) %>%
  summarise(mean_TE = mean(number_of_te, na.rm=T), 
            sd_TE = sd(number_of_te, na.rm=T),
            min_TE = min(number_of_te, na.rm=T),
            max_TE = max(number_of_te, na.rm=T))

TE_byV_anova <- aov(propTE ~ V_subset, data = stacked_byV_wide)
summary(TE_byV_anova)

pairwise.t.test(stacked_byV_wide$number_of_te, stacked_byV_wide$V_subset,
                 p.adjust.method = "BH")

## Anova to compare the unique dom vocab size and unique non-dom vocab size among the 3 developmental groups in the stacked barchart
stacked_byV_wide %>%  
  group_by(BalSubset) %>%
  summarise(mean_Du = mean(Dom_unique_words, na.rm=T), 
            sd_Du = sd(Dom_unique_words, na.rm=T),
            min_Du = min(Dom_unique_words, na.rm=T),
            max_Du = max(Dom_unique_words, na.rm=T))

Du_byB_anova <- aov(Dom_unique_words ~ BalSubset, data = stacked_byV_wide)
summary(Du_byB_anova)

pairwise.t.test(stacked_byV_wide$Dom_unique_words, stacked_byV_wide$BalSubset,
                 p.adjust.method = "BH")

stacked_byV_wide %>% 
  group_by(BalSubset) %>%
  summarise(mean_Nu = mean(NonDom_unique_words, na.rm=T), 
            sd_Nu = sd(NonDom_unique_words, na.rm=T),
            min_Nu = min(NonDom_unique_words, na.rm=T),
            max_Nu = max(NonDom_unique_words, na.rm=T))

Nu_byB_anova <- aov(NonDom_unique_words ~ BalSubset, data = stacked_byV_wide)
summary(Nu_byB_anova)

pairwise.t.test(stacked_byV_wide$NonDom_unique_words, stacked_byV_wide$BalSubset,
                 p.adjust.method = "BH")

# Descriptives: Vocabulary measures
```{r}
## Mean word vocab size
keepers_ws_TE %>% 
  summarize(mean_W = mean(word_vocab, na.rm=T), 
            sd_W = sd(word_vocab, na.rm=T),
            min_W = min(word_vocab, na.rm=T),
            max_W = max(word_vocab, na.rm=T))

## Mean TE size
keepers_ws_TE %>% 
  summarize(mean_TE = mean(number_of_te, na.rm=T), 
            sd_TE = sd(number_of_te, na.rm=T),
            min_TE = min(number_of_te, na.rm=T),
            max_TE = max(number_of_te, na.rm=T))

## Mean unique word size
keepers_ws_TE %>% 
  summarize(mean_Du = mean(Dom_unique_words, na.rm=T), 
            sd_Du = sd(Dom_unique_words, na.rm=T),
            min_Du = min(Dom_unique_words, na.rm=T),
            max_Du = max(Dom_unique_words, na.rm=T))

keepers_ws_TE %>% 
  summarize(mean_Nu = mean(NonDom_unique_words, na.rm=T), 
            sd_Nu = sd(NonDom_unique_words, na.rm=T),
            min_Nu = min(NonDom_unique_words, na.rm=T),
            max_Nu = max(NonDom_unique_words, na.rm=T))

t.test(keepers_ws_TE$Dom_unique_words, keepers_ws_TE$NonDom_unique_words, 
       paired = TRUE) # paired-samples t-test to compare Du and Nu

## Stacked barchart to plot TE, Du, and Nu of individual infants (facet by balance)
stacked <- keepers_ws_TE %>%
  select(baby_id, balance_vocab, word_vocab, number_of_te, Dom_unique_words, NonDom_unique_words) %>%
  # Subset balance_vocab by 3 levels 
  #mutate(BalSubset = cut(balance_vocab, breaks = 3, 
  #                       labels=c("MostBalanced", "MediumBalanced", "LeastBalanced"))) %>%
  mutate(BalSubset = case_when(balance_vocab <= 0.65 ~ "MostBalanced",
                               balance_vocab >= 0.80 ~ "LeastBalanced",
                               TRUE ~ "MediumBalanced")) %>%
  arrange(desc(number_of_te)) %>%
  mutate(TE = number_of_te) %>%
  pivot_longer(-c(baby_id, balance_vocab, BalSubset, word_vocab, TE), 
               names_to = "Type", values_to = "NumberOfWords") %>%
  mutate(Type = factor(Type, levels=c("NonDom_unique_words", "Dom_unique_words", "number_of_te"))) %>%
  mutate(Type = recode(Type, 
                       NonDom_unique_words = "Unique non-dominant word", 
                       Dom_unique_words = "Unique dominant word", 
                       number_of_te = "Translation equivalent"))

LeastBal_plot <- stacked %>%
  filter(BalSubset == "LeastBalanced") %>%
  ggplot(aes(x = reorder(baby_id, -NumberOfWords, sum), y = NumberOfWords)) +
  geom_col(aes(fill = Type)) +
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
  scale_y_continuous(limits = c(0, 1050)) +
  labs(x = "Individual infant", 
       y = "Number of concepts produced",
       title = "Least balanced") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))

MediumBal_plot <- stacked %>%
  filter(BalSubset == "MediumBalanced") %>%
  ggplot(aes(x = reorder(baby_id, -NumberOfWords, sum), y = NumberOfWords)) +
  geom_col(aes(fill = Type)) +
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
  scale_y_continuous(limits = c(0, 1050)) +
  labs(x = "Individual infant", 
       y = "Number of concepts produced",
       title = "Medium balanced") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))

MostBal_plot <- stacked %>%
  filter(BalSubset == "MostBalanced") %>%
  ggplot(aes(x = reorder(baby_id, -NumberOfWords, sum), y = NumberOfWords)) +
  geom_col(aes(fill = Type)) +
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
  scale_y_continuous(limits = c(0, 1050)) +
  labs(x = "Individual infant", 
       y = "Number of concepts produced",
       title = "Most balanced") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))

Bal_StackedPlot <- ggarrange(LeastBal_plot, 
          MediumBal_plot + rremove("ylab"), 
          MostBal_plot + rremove("ylab"),
          ncol = 3, nrow = 1,
          common.legend = TRUE, legend="bottom")

ggsave("Output/StackedPlot_byBalance.png", Bal_StackedPlot,
       width = 10, height = 5)

## Anova to compare the total vocab size among the 3 balance groups in the stacked barchart
stacked_wide <- keepers_ws_TE %>%
  select(baby_id, balance_vocab, word_vocab, concept_vocab, propTE, number_of_te, Dom_unique_words, NonDom_unique_words) %>%
  # Subset balance_vocab by 3 levels 
  mutate(BalSubset = case_when(balance_vocab <= 0.65 ~ "MostBalanced",
                               balance_vocab >= 0.80 ~ "LeastBalanced",
                               TRUE ~ "MediumBalanced"))

W_byB_anova <- aov(word_vocab ~ BalSubset, data = stacked_wide)
summary(W_byB_anova)

pairwise.t.test(stacked_wide$word_vocab, stacked_wide$BalSubset, p.adj = "BH")

## Mean proportion of TE
stacked_wide %>%
  group_by(BalSubset) %>%
  summarize(mean_propTE = mean(propTE, na.rm=T), 
            sd_propTE = sd(propTE, na.rm=T),
            min_propTE = min(propTE, na.rm=T),
            max_propTE = max(propTE, na.rm=T))

## Correlation between concept vocab and TE across different balance group
stacked_wide_most <- stacked_wide %>%
  filter(BalSubset == "MostBalanced")

cor.test(stacked_wide_most$number_of_te, stacked_wide_most$concept_vocab, method = "pearson")

stacked_wide_med <- stacked_wide %>%
  filter(BalSubset == "MediumBalanced")

cor.test(stacked_wide_med$number_of_te, stacked_wide_med$concept_vocab, method = "pearson")

stacked_wide_least <- stacked_wide %>%
  filter(BalSubset == "LeastBalanced")

cor.test(stacked_wide_least$number_of_te, stacked_wide_least$concept_vocab, method = "pearson")

## Anova to compare the unique dom vocab size and unique non-dom vocab size among the 3 balance groups in the stacked barchart
stacked_wide %>% 
  group_by(BalSubset) %>%
  summarize(mean_Du = mean(Dom_unique_words, na.rm=T), 
            sd_Du = sd(Dom_unique_words, na.rm=T),
            min_Du = min(Dom_unique_words, na.rm=T),
            max_Du = max(Dom_unique_words, na.rm=T))

Du_byB_anova <- aov(Dom_unique_words ~ BalSubset, data = stacked_wide)
summary(Du_byB_anova)

pairwise.t.test(stacked_wide$Dom_unique_words, stacked_wide$BalSubset,
                 p.adjust.method = "BH")

stacked_wide %>% 
  group_by(BalSubset) %>%
  summarize(mean_Nu = mean(NonDom_unique_words, na.rm=T), 
            sd_Nu = sd(NonDom_unique_words, na.rm=T),
            min_Nu = min(NonDom_unique_words, na.rm=T),
            max_Nu = max(NonDom_unique_words, na.rm=T))

Nu_byB_anova <- aov(NonDom_unique_words ~ BalSubset, data = stacked_wide)
summary(Nu_byB_anova)

pairwise.t.test(stacked_wide$NonDom_unique_words, stacked_wide$BalSubset,
                 p.adjust.method = "BH")
```

@@

## Merge the two percentile dataframe
ws_percentile <- merge(english_ws_percentile, french_ws_percentile, 
                       by = "age") %>%
  # compute estimate total vocab by adding the two languages
  mutate(Average_90percentile = (EngWS_90percentile + FrWS_90percentile)/2) %>%
  # keep only necessary columns
  select(age, EngWS_90percentile, FrWS_90percentile, Average_90percentile) %>%
  # rename age to prepare for the next step (i.e., merging with keepers_ws_TE
  rename(age_months_percentile = age)

write.csv(ws_percentile, "Output/ws_percentile.csv", row.names = F)
```

# Adding percentile info to keepers_ws_TE dataframe & 
## Creating a new propTE based on the percentile
```{r}
# merge the dataframes
keepers_ws_TE <- keepers_ws_TE %>%
  # as the upper age limit of the CDI-WS is 30m, create a new age variable which replaces all >30m as 30m for merging with the percentile dataframe
  # this means the percentile should be the same for all >30m
  mutate(age_months_percentile = ifelse(age_months_binned > 30, 30, age_months_binned)) %>%
  # merge with the percentile data frame
  left_join(ws_percentile, by = "age_months_percentile")

%>%
  # calculate TE in relation to CDI percentile data
  mutate(propTE_percentile = number_of_te*2/Total_90percentile)
  # %>% select(baby_id, age_months_binned, age_months_percentile, word_vocab, Total_90percentile,
  #       propTE, propTE_CDI, propTE_percentile)

@@

# Correlation
```{r}
Correlation <- keepers_ws_TE %>%
  select(age_continuous, Average_90percentile, balance_vocab, word_vocab, TotalNW_Dom, TotalNW_NonDom,
         number_of_te, Dom_unique_words, NonDom_unique_words, concept_vocab) %>%
  # recode the balance_vocab so that it would be in descending order, so the value of the most balanced would be the greatest
  mutate(balance_vocab = balance_vocab*-1) %>%
  # rename variables
  rename(Age_month = age_continuous,
         V = Average_90percentile,
         B = balance_vocab, 
         W = word_vocab,
         D = TotalNW_Dom, 
         N = TotalNW_NonDom, 
         TE = number_of_te, 
         Du = Dom_unique_words, 
         Nu = NonDom_unique_words, 
         C = concept_vocab)

library(Hmisc)
mycor <- rcorr(as.matrix(Correlation), type="pearson")
mycor$r
round(mycor$P, 2)

library(corrplot)
mycor <- round(cor(Correlation), 2)
corrplot(mycor)
write.csv(mycor, "Output/ObservedTE_correlations.csv", col.names=NA)

cor.test(Correlation$TE, Correlation$B, method = "pearson")
cor.test(Correlation$TE, Correlation$W, method = "pearson")
cor.test(Correlation$TE, Correlation$D, method = "pearson")
cor.test(Correlation$TE, Correlation$N, method = "pearson")
cor.test(Correlation$TE, Correlation$V, method = "pearson")
cor.test(Correlation$Du, Correlation$B, method = "pearson")
cor.test(Correlation$Nu, Correlation$B, method = "pearson")

@@

###### Theoretical distribution of TEs (Simulation): D * N / V (TotalNW_Dom * TotalNW_NonDom / Average_90percentile) ###### 
```{r}
## Create predicted TEs for simulation
keepers_ws_TE_simulation <- keepers_ws_TE %>%
  mutate(Predicted_TE_percentile = TotalNW_Dom * TotalNW_NonDom/Average_90percentile)

## Linear regressions of predicted TE models
Percentile_model <- keepers_ws_TE_simulation %>%
  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
     data = .)

summary(Percentile_model)

## Ploting Y = observed rawTE  & X = Predicted rawTE 
TEprediction_percentile <- keepers_ws_TE_simulation %>%
  ggplot(aes(x = Predicted_TE_percentile, y = number_of_te)) +
#  facet_grid(. ~TEgroup) +
  labs(x = "Number of TEs predicted by the Bilingual Vocabulary Model", y = "Observed total number of TEs") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_smooth(method = "lm", se = F) +
  geom_point(shape = 23) +
  geom_abline(linetype = 2, alpha = 0.5) +
  xlim(-10,430) +
  ylim(-10,430) +
  theme_minimal()

ggsave(here("Output/TEprediction_percentile.png"), TEprediction_percentile,
       height=5, width=5)

# Kolmogorov-Smirnov test to compare the observed and predicted TE
#stats::ks.test(keepers_ws_TE_simulation$number_of_te,
#               keepers_ws_TE_simulation$Predicted_TE_percentile)

## Looking closer at the infants with total vocabulary < 150

keepers_ws_TE_simulation_LessThan300 <- keepers_ws_TE_simulation %>%
  filter(word_vocab < 300)

stats::ks.test(keepers_ws_TE_simulation_LessThan300$number_of_te,
               keepers_ws_TE_simulation_LessThan300$Predicted_TE_percentile)

keepers_ws_TE_simulation_MoreThan300 <- keepers_ws_TE_simulation %>%
  filter(word_vocab > 300)

stats::ks.test(keepers_ws_TE_simulation_MoreThan300$number_of_te,
               keepers_ws_TE_simulation_MoreThan300$Predicted_TE_percentile)

# Plot
TEprediction_percentile_facetW <- keepers_ws_TE_simulation %>%
  mutate(W_group = if_else(word_vocab < 300, "Less than 300 total vocabulary", "More than 300 total vocabulary")) %>%
  ggplot(aes(x = Predicted_TE_percentile, y = number_of_te)) +
  facet_wrap(~ W_group, scales = "free") +
  #facet_grid(. ~ W_group, scales = "free") +
  labs(x = "Number of TEs predicted by the Bilingual Vocabulary Model", y = "Observed total number of TEs") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_smooth(method = "lm", se = F) +
  geom_point(shape = 23) +
  geom_abline(linetype = 2, alpha = 0.5) +
  theme_minimal()

ggsave(here("Output/TEprediction_percentile_facetW.png"), TEprediction_percentile_facetW,
       height=5,width=10)

Percentile_model_LessThan300W <- keepers_ws_TE_simulation %>%
  filter(word_vocab < 300) %>%
  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
     data = .)

summary(Percentile_model_LessThan300W)

Percentile_model_MoreThan300W <- keepers_ws_TE_simulation %>%
  filter(word_vocab > 300) %>%
  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
     data = .)

summary(Percentile_model_MoreThan300W)
